Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-20 docs/manuscript/manuscript.tex
Explanation of figure 4.

Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables. Majority of the data is from 2010 - 2020 (Figure 1). Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys. Most of the variables are related to water accessibility by their locations, time, and water resources. This also allows us to find water accessibility predictor variables and new clustering graphs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{year}
  \caption{Distribution of years of water accessibility variables}
  \label{fig:year}
\end{figure}
\subsection*{Model results}
First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility. We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has strongest and positive relationship with Households possessing a bicycle (bicy). Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws). This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and relationship between these variables.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{correlation-basic}
  \caption{Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p value and to see relationship between different categorical variables is shown in Figure 3, which again indicates greater detailed comparison between the variables. Households with improved water source on the premises (iwsp) has strong relationship with  households with basic water service (basw) of .84 and p value of less than 0.001. Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using an unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with p value of less than 0.001 as well. These relationships justify the necessity to research further if the particular water accessibility variables can increase accessibility of particular water sources in order to reach the universal access to water by 2030, as proposed in the SDGs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{adv-cor-plot}
  \caption{Advanced Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables. Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which is shown in Figure 4,5,6.

>>>>

Across 78 countries in this study, the data provides 28 numerical variables and four descriptive variables. The majority of the data is from 2010 - 2020 (Figure 1). Therefore, we can analyze the global condition of water accessibility by utilizing the most up to date surveys. Most of the variables are related to water accessibility by their locations, time, and water resources. This also allows us to find water accessibility predictor variables and new clustering graphs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{year}
  \caption{Distribution of years of water accessibility variables}
  \label{fig:year}
\end{figure}
\subsection*{Model results}
First, we developed a correlation chart to quickly analyze if there is any great correlation between all the 28 numerical variables that may pertain to water accessibility. We recognized the strong and weak relationship by the size and color of the correlation plot is shown in Figure 2, which indicates households with limited water service (liws) has the strongest and positive relationship with Households possessing a bicycle (bicy). Meanwhile, households using an improved water source (imws) has the strongest and negative relationship with households using an unimproved water source (uiws). This initial finding results indicated that the initial variables are logical and can give us insight into further understanding and the relationship between these variables.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{correlation-basic}
  \caption{Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
As we continued to study further upon the correlations of the variables, we used more advanced correlation graphs to determined their p-value, and to see a relationship between different categorical variables is shown in Figure 3, which again indicates a greater detailed comparison between the variables. Households with an improved water source on the premises (iwsp) have a strong relationship with households with basic water service (basw) of .84 and a p-value of less than 0.001. Also, we learned and discovered that Households using an unimproved water source (uiws) have a positive relationship with Households using unprotected well water (uwel), Households using an unprotected spring (uspr), and Households using surface water (surw) with a p-value of less than 0.001 as well. These relationships justify the necessity to research further if the particular water accessibility variables can increase the accessibility of particular water sources to reach universal access to water by 2030, as proposed in the SDGs.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{adv-cor-plot}
  \caption{Advanced Correlation Graph of Numerical Variables}
  \label{fig:year}
\end{figure}
Additionally, we wanted to explore if the modes of transportation variables have any relationship with potential water variables. Therefore, we investigated by developing more correlation plots relating transportation variables with water accessibility variables which are shown in Figure 4,5,6. There is a significant positive relationship between households possessing a private car (pcar) and households with water on the premises (watp) as shown in Figure 4. There is also is a significant negative relationship between households possessing a private car (pcar) with households with water 30 minutes or less away round trip (tles) and households with water more than 30 minutes away round trip (tmor) with a p-value of less than 0.001. This information is essential to investigate further why this is the global trend.
